\pkg{Rgbp}: An R Package for Gaussian, Poisson, and Binomial Hierarchical Modeling

### 2013-11-12 jss/article.tex
warning added + art changes

\section[introduction]{Introduction}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.
\pkg{Rgbp} is an R package for estimating and validating a two-level model, also called conjugate hierarchical generalized linear model \citep{hglm2006, hglm2010} or called a conditionally independent hierarchical model \citep{1989}, producing point and interval estimates of  random effects. The estimation procedure utilizes approximate Bayesian machinery and the validation involves checking frequency properties of the procedure via repeated sampling (which we call ``frequency method checking"). It is found that even in small samples our procedure yields good frequency properties in comparison to other methods such as Maximum Likelihood Estimation (MLE).

@@

where $\mu_{0j} = x_{j}' \mathbf{\beta}_{(m\times1)}=\beta_{0}+\beta_{1}x_{1, j} + \cdots + \beta_{m-1}x_{m-1, j}$, $j=1, \ldots k$ and $m$ is the number of regression coefficients including an intercept term. The notation $x_{j}'=(1, x_{1, j}, x_{2, j}, \ldots, x_{m-1, j})$ represents the $j$-th group's $m\times 1$ covariate vector whose first element is 1 for the intercept term of regression coefficients. Both $\mathbf{\beta}_{(m\times1)}$ and the second-level variance $A$ are unknown. Note that if there is no covariate then $x_{j}=1$ just for an intercept term ($m=1$) and so $\mu_{0j}=\mu_{0}=\beta_{0}$ for all $j$, resulting in an exchangeable prior distribution. For reference, a parameter with a zero subscript, such as $\mu_{0j}$, represents a mean parameter of the prior (second-level) distribution, \emph{i.e.}, a prior mean. Based on this conjugate prior distribution it is easy to derive the corresponding posterior distribution

@@

The Poisson model can have such posterior propriety if there are two groups whose observed values exceed 0 and the second-level mean ($\lambda_0$) is known. If $\lambda_0$ is unknown, the posterior distribution for the Poisson model is improper and this model should not be used. However, the Binomial model can still be used  instead because the Poisson distribution is just the approximation to the Binomial distribution. The efficiency of these priors is explored in the examples in section 5.

>>>>

\section[introduction]{Introduction}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.
\pkg{Rgbp} is an R package for estimating and validating a two-level model, also known as a conjugate hierarchical generalized linear model \citep{hglm2006, hglm2010} or a conditionally independent hierarchical model \citep{1989}, producing point and interval estimates of  random effects. The estimation procedure utilizes approximate Bayesian machinery and the validation involves checking frequency properties of the procedure via repeated sampling (which we call ``frequency method checking"). It is found that even in small samples our procedure yields good frequency properties in comparison to other methods such as Maximum Likelihood Estimation (MLE).

@@

where $\mu_{0j} = x_{j}' \mathbf{\beta}_{(m\times1)}=\beta_{0}+\beta_{1}x_{1, j} + \cdots + \beta_{m-1}x_{m-1, j}$, $j=1, \ldots k$ and $m$ is the number of regression coefficients including an intercept term. The notation $x_{j}'=(1, x_{1, j}, x_{2, j}, \ldots, x_{m-1, j})$ represents the $j$-th group's $m\times 1$ covariate vector whose first element is 1 for the intercept term of regression coefficients. It is assumed that the second-level variance $A$ is unknown and that $\mathbf{\beta}_{(m\times1)}$ is also unknown unless specified otherwise. Note that if there are no covariates then $x_{j}=1$, ($m=1$) and so $\mu_{0j}=\mu_{0}=\beta_{0}$ for all $j$, resulting in an exchangeable prior distribution. For reference, a parameter with a zero subscript, such as $\mu_{0j}$, represents a mean parameter of the prior (second-level) distribution, \emph{i.e.}, a prior mean. Based on this conjugate prior distribution it is easy to derive the corresponding posterior distribution

@@

The Poisson model can have such posterior propriety if there are two groups whose observed values exceed 0 and the second-level mean ($\lambda_0$) is known. If $\lambda_0$ is unknown, the posterior distribution for the Poisson model is improper and this model should not be used. In the case where the Poisson model is as an approximation to the Binomial and the exposures, $n_j$, are known then the Binomial model should be used as an alternative to the Poisson. The efficiency of these priors is explored in the examples in section 5.
